Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2011-05-08 release_version/supplementary.tex
Moving release materials into one directory

% Include other packages here, before hyperref.
\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

% \usepackage[small]{caption}

@@

\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}

\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}

\begin{document}
%
% paper title
\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
Illumination by Sparse Representation\\
(Supplementary Materials)}
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
        John~Wright,~\IEEEmembership{Member,~IEEE,}
        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
        Hossein~Mobahi,
        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{A. Wagner, A. Ganesh, Z. Zhou, and Y. Ma are with the Dept. of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign. H. Mobahi is with the Computer Science Dept. at the University of Illinois at Urbana-Champaign.  J. Wright and Y. Ma are with Microsoft Research Asia.  Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W. Main st. Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.

\maketitle

\IEEEdisplaynotcompsoctitleabstractindextext
\vspace{-9mm} % HACK

\IEEEpeerreviewmaketitle

%%%%%%%%% BODY TEXT
\section{L1 Minimization via Augmented Lagrange Multiplier}

\PARstart{I}{n} these supplementary materials we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}, which is
repeated here for convenience. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.
\input{deformable-src.tex}

@@

In order to discuss the solution to the above subproblems, we
need to define the following soft-thresholding operator for a
vector $\x$ and a scalar $\alpha \geq 0$:
\begin{equation}
\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
\end{equation}
where all the operations are performed component-wise. It is
easy to show that the subproblem with respect to $\e$ has a
closed-form solution given by the soft-thresholding operator:
\begin{equation}
\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
\end{equation}
To solve the subproblem associated with $\x$, we
apply a first-order $\ell^1$-minimization method,
called \emph{fast iterative shrinkage-threshold algorithm}
(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
iteratively minimize a quadratic approximation $Q(\x, \z)$ to
$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
carefully chosen in order to achieve a good convergence
rate. We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
\begin{algorithm}[t]
\caption{\bf (Augmented Lagrange Multiplier Method for Global
Recognition)}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\label{alg:alm}
%\vspace{-4mm}
\end{algorithm}

We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory,
running on our database containing images size $80\times 60$
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

@@

\input{bios/bios}

>>>>

\usepackage{verbatim}

\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

@@

\begin{document}

\appendix[L1 Minimization via Augmented Lagrange Multiplier]
In this Appendix we discuss the computational issues related to
the implementation of Algorithm 1, which is
repeated here for convenience. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.

@@

In order to discuss the solution to the above subproblems, we
need to define the following soft-thresholding operator for a
vector $\x$ and a scalar $\alpha \geq 0$:
\begin{equation}
\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
\end{equation}
where all the operations are performed component-wise. It is
easy to show that the subproblem with respect to $\e$ has a
closed-form solution given by the soft-thresholding operator:
\begin{equation}
\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
\end{equation}
To solve the subproblem associated with $\x$, we
apply a first-order $\ell^1$-minimization method,
called \emph{fast iterative shrinkage-threshold algorithm}
(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
iteratively minimize a quadratic approximation $Q(\x, \z)$ to
$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
carefully chosen in order to achieve a good convergence
rate. We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
\begin{algorithm}[t]
\caption{\bf (Augmented Lagrange Multiplier Method for Global
Recognition)}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\label{alg:alm}
\end{algorithm}

We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory,
running on our database containing images size $80\times 60$
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm 1 takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

@@
